---
title: "Endothelial GD expression"
author: "Ruiyu Ray Wang"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
wd = '/Users/Ray/Documents/Work/scRNAseq/EC & Single-Cell Sequencing'
READ_CACHE = TRUE
```
```{r}
library(Seurat)
library(org.Mm.eg.db)
library(scater)
library(ggplot2)
library(RColorBrewer)
library(awesomeSC)
```

## LPS condition
### Duan et al. 2018
```{r}
if (READ_CACHE){
  cells.int <- readRDS(file = file.path(wd, "LPS Treatment", "Duan 2018", "cells.int.annotated.rds"))
} else {
  SA01 <- read.csv(file = file.path(wd, "LPS Treatment", "Dxuan 2018", "GSM3070094_SA01.csv.gz"), row.names = 1)
  cells <- CreateSeuratObject(SA01); cells$treat <- "Saline"; rm(SA01)
  SA02 <- read.csv(file = file.path(wd, "LPS Treatment", "Duan 2018", "GSM3070095_SA02.csv.gz"), row.names = 1)
  cells.SA02 <- CreateSeuratObject(SA02); cells.SA02$treat <- "Saline"; cells <- merge(cells, cells.SA02); rm(SA02, cells.SA02)
  SA03 <- read.csv(file = file.path(wd, "LPS Treatment", "Duan 2018", "GSM3070096_SA03.csv.gz"), row.names = 1)
  cells.SA03 <- CreateSeuratObject(SA03); cells.SA03$treat <- "Saline"; cells <- merge(cells, cells.SA03); rm(SA03, cells.SA03)
  LPS01 <- read.csv(file = file.path(wd, "LPS Treatment", "Duan 2018", "GSM3070097_LPS01.csv.gz"), row.names = 1)
  cells.LPS01 <- CreateSeuratObject(LPS01); cells.LPS01$treat <- "LPS"; cells <- merge(cells, cells.LPS01); rm(LPS01, cells.LPS01)
  LPS02 <- read.csv(file = file.path(wd, "LPS Treatment", "Duan 2018", "GSM3070098_LPS02.csv.gz"), row.names = 1)
  cells.LPS02 <- CreateSeuratObject(LPS02); cells.LPS02$treat <- "LPS"; cells <- merge(cells, cells.LPS02); rm(LPS02, cells.LPS02)
  LPS03 <- read.csv(file = file.path(wd, "LPS Treatment", "Duan 2018", "GSM3070099_LPS03.csv.gz"), row.names = 1)
  cells.LPS03 <- CreateSeuratObject(LPS03); cells.LPS03$treat <- "LPS"; cells <- merge(cells, cells.LPS03); rm(LPS03, cells.LPS03)
}
```

### QC by scater

Convert to SingleCellExperiment object, convert row names, and run scater QC pipeline.
```{r}
cells.sce <- as.SingleCellExperiment(cells)

# Rename rows from ENSEMBL ID to gene SYMBOL
gn <- mapIds(x = org.Mm.eg.db, keys = rownames(cells.sce), column = "SYMBOL", keytype = "ENSEMBL", multiVals = "first")  # 5606 NAs produced
# anno <- loadGencodeAnno()
# rn <- rownames(cells.sce)
# gn <- anno$gene_short_name[match(rn, anno$stable_id)]  # 276 NAs produced
rownames(cells.sce) <- gn

# Prune rows with NA names
cells.sce <- cells.sce[!is.na(rownames(cells.sce)),]
```

Perform QC steps.
```{r}
# Remove unexpressed genes to save memory and computation time
keep_feature <- rowSums(as.matrix(counts(cells.sce)) > 0) > 0
cells.sce <- cells.sce[keep_feature, ]
```

Cell level QC.
```{r}
per.cell <- perCellQCMetrics(cells.sce)
summary(per.cell$sum)
summary(per.cell$detected)
colData(cells.sce) <- cbind(colData(cells.sce), per.cell)  # store to colData()
```
```{r quickPerCellQC}
do_quickPerCellQC = TRUE

if (do_quickPerCellQC){
  qc.stats <- quickPerCellQC(per.cell)
  colSums(as.matrix(qc.stats))
}
cells.filtered <- cells.sce[,!qc.stats$discard]
```

Feature-level QC.
```{r perFeatureQCMetrics}
per.feat <- perFeatureQCMetrics(cells.filtered)
summary(per.feat$mean)
summary(per.feat$detected)
```
```{r filtering features}
keep_feature <- nexprs(cells.filtered, byrow=TRUE) > 0
cells.filtered <- cells.filtered[keep_feature,]
dim(cells.filtered)
```

Variable QC: omitted.

### Seurat pipeline

Convert from SCE object to Seurat obejct.
```{r}
cells <- as.Seurat(cells.filtered)
```

Run integration pipeline.
```{r}
# Split object by treatment: Saline or LPS
cells.list <- SplitObject(cells, split.by = "treat")
# Iteratively perform Normalization and Feature Selection
cells.list <- lapply(X = cells.list, FUN = function(x){
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})
# Find anchors
cells.anchors <- FindIntegrationAnchors(cells.list, dims = 1:30)
# Perform integration
cells.int <- IntegrateData(anchorset = cells.anchors, dims = 1:30)
```


```{r fig.width=12, fig.height=5}
cells.int <- SSW(cells.int, assay = "integrated", npcs = 30, dims = 1:30, verbose = FALSE)
cells.int <- SSW(cells.int, assay = "RNA", npcs = 30, dims = 1:30, verbose = FALSE)

# Make UMAP look prettier
cells.int <- RunUMAP(cells.int, reduction = "pca", dims = 1:30, umap.method = 'umap-learn', metric = 'correlation', n.epochs = 400, reduction.key = "UMAP_", reduction.name = "umap", verbose = FALSE)
cells.int <- RunTSNE(cells.int, reduction = "pca", dims = 1:30, seed.use = 2, tsne.method = "Rtsne", reduction.key = "TSNE_", reduction.name = "tsne", verbose = FALSE)

DimPlot(cells.int, group.by = "RNA_snn_res.0.3", reduction = "tsne", label = TRUE) | DimPlot(cells.int, group.by = "treat", reduction = "tsne")
```

Multi-gene feature plot.
```{r}
genes_to_plot <- c("Gfap",  # Astrocyte
                   "Pecam1",  # Endothelia
                   "Mobp", "Bcas1", "C1ql1", "Enpp2",  # Oligodendrocytes
                   "Map2", "Rbfox3", "Nrgn", "Wfs1", "Spock1", "Reln", "Gad2",  # Neurons
                   "Aif1", "Cx3cr1",  # Microglia
                   "Mrc1"  # Macrophage
                   )
FeaturePlot(cells.int, features = genes_to_plot, reduction = "tsne")
```

The cell identities were annotated by referencing the original paper (Duan et al. 2018), and cross-validating with online resources and databases (mousebrain.org).

```{r fig.width=12, fig.height=5}
# Save/Load manually identified cells
save(opcs,rbcs,peri,miscellaneous,endo.keep, file = file.path(wd,"LPS Treatment", "Duan 2018", "aux_ids.dat"))
load(file.path(wd,"LPS Treatment", "Duan 2018", "aux_ids.dat"))
# Detailed identity with gene markers
Idents(cells.int) <- "RNA_snn_res.0.3"
cells.int <- RenameIdents(cells.int, `0` = "Astrocytes", `1` = "Nrgn Neurons", `2` = "Immature Neurons", `3` = "Endothelial cells", 
                          `4` = "Mobp Oligodendrocytes", `5` = "Endothelial cells", `6` = "C1ql1 Oligodendrocytes", 
                          `7` = "Neuronal intermediate progenitor cells", `8` = "Choiroid plexus epithelial cells", `9` = "Bcas1 Oligodendrocytes",
                          `10` = "Wfs1 Neurons", `11` = "Cajal-Retzius cells", `12` = "Microglia", `13` = "Pericytes", `14` = "Interneurons", 
                          `15` = "Tfap2c Astrocytes", `16` = "Astrocytes", `17` = "Vascular Leptomeningeal cells", `18` = "Spock1 Neurons", 
                          `19` = "Spock1 Neurons", `20` = "Ependymal cells", `21` = "Macrophages", `22` = "Neurons", `23` = "Neurons")

cells.int <- SetIdent(cells.int, cells = opcs, value = "Oligodendrocytes precursor cells")
cells.int <- SetIdent(cells.int, cells = rbcs, value = "Red Blood Cells")
cells.int <- SetIdent(cells.int, cells = peri, value = "Pericytes")
cells.int$cell_type_detailed <- Idents(cells.int)

# General identity
Idents(cells.int) <- "RNA_snn_res.0.3"
cells.int <- RenameIdents(cells.int, `0` = "Astrocytes", `1` = "Neurons", `2` = "Neurons", `3` = "Endothelial cells", 
                          `4` = "Oligodendrocytes", `5` = "Endothelial cells", `6` = "Oligodendrocytes", `7` = "Neurons", 
                          `8` = "Choiroid plexus epithelial cells", `9` = "Oligodendrocytes", `10` = "Neurons", `11` = "Cajal-Retzius cells", 
                          `12` = "Microglia", `13` = "Pericytes", `14` = "Interneurons", `15` = "Astrocytes", `16` = "Astrocytes",
                          `17` = "Vascular Leptomeningeal cells", `18` = "Neurons", `19` = "Neurons", `20` = "Ependymal cells", 
                          `21` = "Macrophages", `22` = "Neurons", `23` = "Neurons")
cells.int <- SetIdent(cells.int, cells = opcs, value = "Oligodendrocytes precursor cells")
cells.int <- SetIdent(cells.int, cells = rbcs, value = "Red Blood Cells")
cells.int <- SetIdent(cells.int, cells = peri, value = "Pericytes")
cells.int$cell_type <- Idents(cells.int)

# Abbreviated identity
Idents(cells.int) <- "RNA_snn_res.0.3"
cells.int <- RenameIdents(cells.int, `0` = "Astro", `1` = "Neurons", `2` = "Neurons", `3` = "Endo", 
                          `4` = "OL", `5` = "Endo", `6` = "OL", `7` = "Neurons", 
                          `8` = "CPEpiC", `9` = "OL", `10` = "Neurons", `11` = "CR", 
                          `12` = "MG", `13` = "Peri", `14` = "Interneurons", `15` = "Astro", `16` = "Astro",
                          `17` = "VLMC", `18` = "Neurons", `19` = "Neurons", `20` = "Epen", 
                          `21` = "MΦ", `22` = "Neurons", `23` = "Neurons")
cells.int <- SetIdent(cells.int, cells = opcs, value = "OPCs")
cells.int <- SetIdent(cells.int, cells = rbcs, value = "RBCs")
cells.int <- SetIdent(cells.int, cells = peri, value = "Peri")
cells.int$cell_type_brief <- Idents(cells.int)

# library(colorspace)
# pal <- choose_palette()
display.brewer.pal(n = 12, name = "Paired")
cols <- brewer.pal(n = 12, name = "Paired")
DimPlot(cells.int, group.by = "cell_type_brief", reduction = "tsne", label = TRUE) | DimPlot(cells.int, group.by = "treat", reduction = "tsne", cols = cols[c(10,7)])
```

Backup.
```{r}
saveRDS(cells.int, file = file.path(wd, "LPS Treatment", "Duan 2018", "cells.int.annotated.rds"))
```

#### FigX.A

```{r}
dA1_left <- DimPlot(cells.int, group.by = "cell_type_brief", reduction = "tsne.int") + 
  NoLegend() +
  NoAxes()
ggsave(filename = "Fig_A_int.eps", plot = dA1_left, device = "eps", path = "~/Documents/Work/scRNAseq/EC & Single-Cell Sequencing/LPS Treatment/Duan 2018/figs/A/", dpi = "print", width = 6, height = 5)

dA1_right <- DimPlot(cells.int, group.by = "treat", reduction = "tsne.int", cols = cols[c(10,7)]) + 
  NoLegend() +
  NoAxes()
ggsave(filename = "Fig_A_right.eps", plot = dA1_right, device = "eps", path = "~/Documents/Work/scRNAseq/EC & Single-Cell Sequencing/LPS Treatment/Duan 2018/figs/A/", dpi = "print", width = 6, height = 5)
```
```{r}
f_gd <- FeaturePlot(cells.int, features = "Gsdmd", reduction = "tsne.int") + NoAxes()
```

Subsetting out Endothelial cells for closer examination.
```{r}
if (READ_CACHE){
  cells.endo <- readRDS(file.path(wd, "LPS Treatment", "Duan 2018", "cells.endo.rds"))
} else {
  cells.endo <- subset(cells.int, cells = endo.keep)
}
# Some Astrocytes and Neurons were erroneously mixed into this group, and were removed by manually curating cell ids.
```

```{r}
cells.endo <- SSW(cells.endo, assay = "RNA", npcs = 30, dims = 1:30, verbose = FALSE)
for(i in seq(400,800, by = 10)){
  cells.endo <- RunUMAP(cells.endo, dims = 1:30, n.epochs = i, umap.method = "umap-learn")
  dplot <- DimPlot(cells.endo, group.by = "treat") + NoLegend() + NoAxes()
  ggsave(filename = paste0("dplot_",i,"_epochs.jpeg"), plot = dplot, device = "jpeg", path = "LPS Treatment/Duan 2018/figs/test_epoch/", width = 4, height = 4)
}

for(i in seq(1,30)){
  cells.endo <- RunTSNE(cells.endo, dims = 1:30, seed.use = i)
  dplot <- DimPlot(cells.endo, reduction = "tsne", group.by = "treat") + NoLegend() + NoAxes()
  ggsave(filename = paste0("dplot_seed_",i,".jpeg"), plot = dplot, device = "jpeg", path = "LPS Treatment/Duan 2018/figs/test_epoch/", width = 4, height = 4)
}  # t-SNE with seed = 20 seems to give best look figure

cells.endo <- RunTSNE(cells.endo, dims = 1:30, seed.use = 20)
d_plot_use <- DimPlot(cells.endo, group.by = "treat", reduction = "tsne", cols = cols[c(10,7)]) + NoLegend() + NoAxes(); d_plot_use
out_dir <- '~/Documents/Work/scRNAseq/EC & Single-Cell Sequencing/LPS Treatment/Duan 2018/figs/'
ggsave(filename = "dplot_endo.eps", device = "eps", path = out_dir, width = 4, height = 4, dpi = 300, plot = d_plot_use)
```

```{r}
out_dir <- '~/Documents/Work/scRNAseq/EC & Single-Cell Sequencing/LPS Treatment/Duan 2018/figs/goi_featureplots'
if(!dir.exists(out_dir)){dir.create(out_dir)}
genes_of_interest <- c("Casp1","Casp4","Gsdmd","Tlr4","Tlr3","Cd14","Lbp","Nlrp3","Pycard","Il1b","Tnf","Il6")

iter_feature_plot_save <- function(object, features_plot, reduction = "umap.int", brewer.use = "RdPu", out_dir){
  for (feature in features_plot){
    # Original FeaturePlot
    p <- FeaturePlot(object, features = feature, reduction = reduction) +
      scale_colour_gradientn(colours = brewer.pal(n = 5, name = brewer.use))
    # FeaturePlot without legend, title, axes.
    p_clean <- FeaturePlot(object, features = feature, reduction = reduction) +
      ggtitle("") + 
      NoAxes() + 
      NoLegend() +
      theme(panel.border = element_rect(colour = "black", fill=NA, size=.5)) + 
      scale_colour_gradientn(colours = brewer.pal(n = 5, name = brewer.use))
    ggsave(filename = paste0("featureplot_",feature,"_",reduction,".eps"), plot = p, device = "eps", path = out_dir, width = 4, height = 4)
    ggsave(filename = paste0("featureplot_",feature,"_",reduction,"_clean.eps"), plot = p_clean, device = "eps", path = out_dir,
           width = 4, height = 4)
  }
}

iter_feature_plot_save(cells.endo, features_plot = genes_of_interest, reduction = "tsne", out_dir = out_dir)
```

```{r}
FeaturePlot(cells.int, features = c("Gsdmd"), reduction = "tsne") + NoAxes()
```

Attempting to annotate cell identities using Garnett. **DO NOT RUN.**
```{r}
# library(garnett)
# classifier <- readRDS("/Users/Ray/Documents/Work/scRNAseq/renjing/Injection Data/reanalyze2/mmBrain_20191017.rds")
# 
# # Construct CellDataSet object
# pd <- AnnotatedDataFrame(cells.int@meta.data)
# fd <- AnnotatedDataFrame(data.frame(row.names = rownames(cells.int),
#                                     gene_short_name = rownames(cells.int)))
# cds <- newCellDataSet(cellData = as.matrix(cells.int@assays$RNA@counts),
#                       phenoData = pd, 
#                       featureData = fd)
# cds <- estimateSizeFactors(cds)
# 
# # Predict cell-type using pre-trained classifier
# cds <- classify_cells(cds, classifier,
#                       db = org.Mm.eg.db,
#                       cluster_extend = TRUE,
#                       cds_gene_id_type = "SYMBOL")  # The results produced by Garnett has reproducibility issues
# 
# table(pData(cds)$cluster_ext_type)
# cells.int <- AddMetaData(object = cells.int, metadata = pData(cds))
# Idents(cells.int) <- "cluster_ext_type"
# cells.int[["cluster_ext_type_brief"]] <- Idents(cells.int)
```










